---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


# Overview

sc0 is a super stripped down core for silicate. All it has is the SC0 model, which is

* `coord`, all coordinates from the input
* `segment`, the .vx0 and .vx1 index of every input segment (row of coord)
* `geometry` map (the record of any paths, `cumsum(nrow)` is a count-index into coord)
* `data` (the feature table)

Compare with silicate: 

```{r}
library(silicate)
library(silicore)
sc_coord(minimal_mesh)
SC0(minimal_mesh)$coord
```

```{r}
sc_edge(minimal_mesh)
SC0(minimal_mesh)$segment
```

```{r}
sc_path(minimal_mesh)
SC0(minimal_mesh)$geometry
```

```{r}
sc_object(minimal_mesh)
SC0(minimal_mesh)$data
```


Performance is good. 

```{r}
rbenchmark::benchmark(SC0(minimal_mesh), 
                      SC(minimal_mesh))
```


```{r}
rbenchmark::benchmark(SC0(inlandwaters), 
                      SC(inlandwaters), replications = 10)
```

For now we are ignoring

* relational indexes
* geometric normalization (might be x/y, x/y/z, other spaces ...)
* any explicitly sequential storage, i.e. ARC and PATH probably will be rebuilt completely


We import the internal `silicate:::get_projection.sf` and for now rely on `gibble` to give the mapping between coordinates and paths. 


This is being tried because silicate was originally defined around PATH, and even
SC was defined in terms of it - unnecessarily. Once [we tried](http://rpubs.com/cyclemumner/367272) a purely tidyverse approach to the problem of decomposing paths to edges it seemed natural to use that as a basis. It means that we can keep the run-length index (gibble) as a way of storing higher level groupings, including holes and multipolygons *optionally*. Here we don't need to encode the actual sequence of coordinates along paths or arcs, because i) it's implicit in geometry/coord tables prior to vertex de-duplication and ii) we might discard them favouring edge-traversal as a way of reconstructing sequences. 


Can we get sense out of it? 

```{r}
x <- SC0(minimal_mesh)
library(ggplot2)
library(dplyr)
tab <- tibble(xs = x$coord$x_[x$segment$.vx0], ys = x$coord$y_[x$segment$.vx0], 
              xend = x$coord$x_[x$segment$.vx1], yend = x$coord$y_[x$segment$.vx1])
ggplot(tab, aes(xs, ys, xend  =xend, yend = yend)) + geom_segment()

x <- SC0(inlandwaters)
## the  sixth object
tst <- rep(x$geometry$object, x$geometry$nrow) == 4
idx <- which(tst)
g <-  rep(x$geometry$path, x$geometry$nrow)
segment <- x$segment %>% dplyr::filter(.vx0 %in% idx & .vx1 %in% idx)
tab <- tibble(xs = x$coord$x_[segment$.vx0], ys = x$coord$y_[segment$.vx0], 
              xend = x$coord$x_[segment$.vx1], yend = x$coord$y_[segment$.vx1], 
              path = g[segment$.vx0])
ggplot(tab, aes(xs, ys, xend  =xend, yend = yend, colour = factor(path))) + geom_segment() + guides(colour = FALSE)


x <- SC0(dodgr::hampi)
g <-  rep(x$geometry$path, x$geometry$nrow)
tab <- tibble(xs = x$coord$x_[x$segment$.vx0], ys = x$coord$y_[x$segment$.vx0], 
              xend = x$coord$x_[x$segment$.vx1], yend = x$coord$y_[x$segment$.vx1], 
              path = g[x$segment$.vx0])
ggplot(tab, aes(xs, ys, xend  =xend, yend = yend, colour = path)) + geom_segment()

```


If anyone can come up with a better name than `gibble` or `geometry` or `geometry map`
for that thing, I'll be really grateful. 